---
title: "ACNC 2020 analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, include = FALSE}
# Load packages, filter data to just env
source('process_data.R')
library(dplyr)
library(tidyr)
library(purrr)
library(car)
df <- years[['2019']] 
```


```{r, include = FALSE}
env_df <- df %>% filter(main.activity == 'Environmental activities')
soc_df <- df19 %>% filter(main.activity == 'Social services')
```

```{r}
df19 %>%
  group_by(main.activity) %>%
  filter(main.activity %in% unique(years[['2017']]$main.activity)) %>%
  slice_max(order_by = total.revenue, n = 5) %>%
  summarise(top5_revenue = mean(total.revenue), top5_FTEs = mean(total.full.time.equivalent.staff), n = n()) %>%
  arrange(desc(top5_revenue)) %>%
  # View()
  write.csv('top5sectors.csv')
```



Number of charities, number of environmental charities

There were `r nrow(df)` charities reporting in 2018 of which `r nrow(env_df)` were environmental charities.

Revenue for all charities was `r sum(df$total.revenue)` of which `r sum(env_df$total.revenue)` went to environmental organisations.

Donations and bequests revenue for all charities was `r sum(df$donations.and.bequests)` of which `r sum(env_df$donations.and.bequests)` went to environmental organisations.


**Revenue breakdown**

```{r}
env_df %>%
  select(donations.and.bequests, revenue.from.investments, revenue.from.government, revenue.from.goods.and.services, all.other.revenue) %>%
  gather(key = 'revenue_source', value = 'amount') %>%
  group_by(revenue_source) %>%
  summarise(total = sum(amount)) %>%
  mutate(prop = prop.table(total),
         midpoint = cumsum(prop) - prop / 2) %>%
  ggplot(aes(x = 1, y = prop, fill = revenue_source)) +
    geom_col(position = 'stack') +
    geom_text(aes(label = scales::percent(round(prop, 4)), size=5),
              position = position_stack(vjust = 0.5))
```

For comparison

```{r}
df19 %>%
  select(donations.and.bequests, revenue.from.investments, revenue.from.government, revenue.from.goods.and.services, all.other.revenue) %>%
  gather(key = 'revenue_source', value = 'amount') %>%
  group_by(revenue_source) %>%
  summarise(total = sum(amount)) %>%
  mutate(prop = prop.table(total),
         midpoint = cumsum(prop) - prop / 2) %>%
  ggplot(aes(x = 1, y = prop, fill = revenue_source)) +
    geom_col(position = 'stack') +
    geom_text(aes(label = scales::percent(round(prop, 4)), size=5),
              position = position_stack(vjust = 0.5))
```

```{r}
sum(env_df$donations.and.bequests, na.rm = TRUE) / sum(df$donations.and.bequests, na.rm = TRUE)

sum(env_df$total.revenue, na.rm = TRUE) / sum(df$total.revenue, na.rm = TRUE)

sum(env_df$total.revenue, na.rm = TRUE)
```



```{r}
nrow(env_df)

round(sum(env_df$total.revenue) / 1000000)

sum(env_df$revenue.from.government) / sum(env_df$total.revenue)

sum(env_df$revenue.from.government) / sum(env_df$total.revenue)
```

The total amount of revenue reported by the `r nrow(env_df)` environmental charities in 2018 was `r round(sum(env_df$total.revenue)/1000000)` million. Most of the funding received by environmental charities came from two sources:
 Government grants (`r round(sum(env_df$revenue.from.government)/sum(env_df$total.revenue) * 100)`% of all revenue); and
 Donations and bequests (`r round(sum(env_df$donations.and.bequests)/sum(env_df$total.revenue) * 100)`% of all revenue).
Fewer organisations received government grants
than donations:
 `r round(nrow(env_df[env_df$donations.and.bequests > 0,])/nrow(env_df) * 100)`% of organisations received at least one donation.
 `r round(nrow(env_df[env_df$revenue.from.government > 0,])/nrow(env_df) * 100)`% of organisations got some funding from government
 

Most environmental charities (`r nrow(env_df %>% filter(charity.size == 'Small'))`) meet the small category and are at the “low end” of the revenue spectrum. The median small organisation had just $`r env_df %>% filter(charity.size == 'Small') %>% select(total.revenue) %>% pull() %>% median(na.rm = TRUE)` in total revenue for the year. These organisations are distinctive not just in the amount of revenue they receive, but the way they operate. Most are volunteer run organisations; 
`r round(nrow(env_df[env_df$total.full.time.equivalent.staff == 0,])/nrow(env_df) * 100)` % of small organisations had no paid staff.

There are `r nrow(env_df %>% filter(charity.size == 'Large'))` large organisations. Together they received `r (env_df %>% filter(charity.size == 'Large') %>% select(total.revenue) %>% sum()) / (env_df %>% select(total.revenue) %>% sum())`% of all environmental charity revenue. The 10 largest organisations account for `r ((env_df %>% arrange(desc(total.revenue)) %>% head(10) %>% select(total.revenue) %>% sum()) / (env_df %>% select(total.revenue) %>% sum()) %>% round()) * 100`% of all revenue. The 30 largest organisations receive more funding than the remaining `r nrow(env_df)`.

```{r}
# Percent covered

env_df %>%
  select(donations.and.bequests) %>%
  arrange(desc(donations.and.bequests)) %>%
  mutate(cumulative = cumsum(donations.and.bequests),
         cumperc = cumulative / sum(donations.and.bequests))

env_df %>%
  select(total.revenue) %>%
  arrange(desc(total.revenue)) %>%
  mutate(cumulative = cumsum(total.revenue),
         cumperc = cumulative / sum(total.revenue))
```


```{r}
env_df %>%
  select(charity.size) %>%
  mutate(charity_size = car::recode(charity.size, "'Small' = 'Small (under $250,000)'; 'Medium' = 'Medium ($250,000 to 1 million)'; 'Large' = 'Large (over $1 million)'")) %>%
  group_by(charity_size) %>%
  count() %>%
  ggplot(aes(x = 1, y = n, fill = charity_size)) +
    geom_col(position = 'stack') +
    geom_text(aes(label = n, size=5),
              position = position_stack(vjust = 0.5))
```

```{r}
env_df %>%
  select(charity.size, total.revenue) %>%
  mutate(charity_size = car::recode(charity.size, "'Small' = 'Small (under $250,000)'; 'Medium' = 'Medium ($250,000 to 1 million)'; 'Large' = 'Large (over $1 million)'")) %>%
  group_by(charity_size) %>%
  summarise(total = sum(total.revenue)) %>%
  mutate(prop = prop.table(total)) %>%
  ggplot(aes(x = 1, y = prop, fill = charity_size)) +
    geom_col(position = 'stack') +
    geom_text(aes(label = scales::percent(round(prop, 4)), size=5),
              position = position_stack(vjust = 0.5))
```


## Giving ##

```{r}
# env_year_sums$year <- c(2014, 2015, 2016, 2017, 2018)

years_figs <- all_years_data %>%
  filter(main.activity == 'Environmental activities') %>%
  group_by(year) %>%
  summarise(total_donations = sum(donations.and.bequests, na.rm = T))
  

years_figs
```

```{r}
years_figs %>%
  ggplot(aes(x = year, y = total_donations)) +
    geom_col() +
    geom_smooth() +
    geom_text(aes(label=total_donations), position=position_dodge(width=0.9), vjust=-0.25)
```



We have seen a small but steady increase in giving to environmental organisations from 170M in 2014 to 266M in 2018. This is in line with increasing total revenue for environmental charities. In fact as a portion of total revenue donations and bequests have remained stable at around 30%.

```{r}
all_years_data %>%
  filter(main.activity == 'Environmental activities') %>%
  mutate(perc_donations = (donations.and.bequests / total.revenue) * 100) %>%
  group_by(year) %>%
  summarise(perc_donations = mean(perc_donations, na.rm = T)) %>%
  filter(!is.infinite(perc_donations)) %>%
  ggplot(aes(x = year, y = perc_donations)) +
    geom_line() +
    expand_limits(y = 0)
```
# Top charities

```{r}
env_df_sorted <- env_df %>% arrange(desc(donations.and.bequests))
env_df_sorted[1:10,] %>% select(charity.name, donations.and.bequests)
```

# Most normal charities

```{r}
env_df_sorted <- env_df %>% arrange(desc(donations.and.bequests))
env_df_sorted[474:483,] %>% select(charity.name, donations.and.bequests)
```

```{r}
df19 %>%
  mutate(decile_rank = ntile(donations.and.bequests,10)) %>%
  group_by(decile_rank) %>%
  summarise(decile_sum = sum(donations.and.bequests)) %>%
  mutate(decile_perc = decile_sum / sum(.$decile_sum))
```

What are the big trends? Climate? Disruption?

```{r}
# Top 20 orgs

all_years_data %>%
  filter(main.activity == 'Environmental activities', year %in% c('2016', '2017', '2018', '2019')) %>%
  group_by(abn) %>%
  summarise(total_donations = sum(donations.and.bequests)) %>%
  left_join(env_df %>%
              select(charity.name, abn, donations.and.bequests)
              ) %>%
  arrange(desc(total_donations)) #%>%
  # write_csv('topx.csv')
```

```{r}

df %>%
  group_by(main.activity) %>%
  summarise(total_donations = sum(donations.and.bequests)) %>%
  mutate(perc_donations = total_donations / sum(.$total_donations))

```
